---
title: "R: Basic Operations"
output: html_notebook
---

# R: Basic Operations

## Comments

Comments in R are denoted by `#`

```{r}
# This is a comment
```

There is no built-in way to do multi-line comments.

## Mathematical Operations

Mathematical operations in R are handled using the standard 4 symbols `+, -, /, and *` . Operations are calculated using PEMDAS.

### Addition

```{r}
# Addition
10 + 512

```

### Subtraction

```{r}
# Subtraction
921 - 14
```

### Multiplication

```{r}
# Multiplication
12 * 12
```

### Division

```{r}
# Division
144  / 12 
392 / 5
```

## Variable Assignment

Variables in R are traditionally assigned using an arrow sign `<-` , made using a less than carat and a dash. It is valid to use `=` , but not recommended.

```{r}
# mac shortcut: option -
x <- 34
x
```

Variable names can not include and spaces or symbols other than the underscore `_` . They cannot begin with numbers, but can include numbers after the first character.

```{r}
this_is_a_valid_name_10 <- "Bob"
this_is_a_valid_name_10
```

## Data Types

R has the following data types: numeric, integer, character, logical, and complex.

### Numeric

Decimal values in R are numerics. Numerics are the default data type for numbers.

```{r}
a = 1.5
a
class(a)

```

```{r}
is.numeric(a)

```

Note: Numerics are not integers. You can confirm this with both a and b above using the `is.integer()` function.

```{r}
is.integer(a)
```

```{r}

```

### Integer

To create an integer in R we can use the `as.integer()` function.

```{r}
i <- as.integer(5)
i
```

```{r}
class(i)
```

For shorthand, we can also define an integer by writing the desired integer value, followed by a capital L.

```{r}
y <- 6L
y
```

```{r}
is.integer(y)
```

Integers must be whole numbers, if passed a floating value, the number will be rounded to an integer.

```{r}
t = as.integer(3.89)
t
```

This also works with a string.

```{r}
intstr <- as.integer("1.8")
intstr
```

We would use an integer data-type when we want to represent a whole integer, and don't want to accept any other variations of numbers. The numeric data-type is less strict, it allows us to switch between different number types like integers and floats dynamically.

### Character

A **character** type is used to represent strings in R. Characters can be denoted by writing within quotes.

```{r}
# Defining a character object
my_str <- 'Jamba'
my_str
```

#### Concatenating character objects

Character objects can be concatenated into a single variable using the `paste` function

```{r}
my_other_str <- 'is delicious!'
message = paste(my_str, my_other_str)
message # note automatically adds space
```

#### Converting into character objects

We can convert other objects into character objects using the `as.character()` function

```{r}
x = as.character(12)
x
```

```{r}
# Note that class returns the data type of our object
class(x)
```

#### Printing character types

You can print out hardcoded character objects with the `print` function

```{r}
print('I want a cookie')
```

You can print out formatted character objects with the `sprintf` function. You can use the following symbols to represent formatted parts of the code, each should be preceded by a `%` symbol:

-   d and i - Integer values

-   s - Character string

-   o - Octal

-   x and X - Hexadecimal

-   f - Double precision value (For fixed point representation, for example: 21.234)

-   e and E - Double precision value (For exponential representation, for example: 2.1234e+01)

```{r}
sprintf('%s are %d cookies.', 'John', 21)
```

```{r}
sprintf('%f %e', 21.234, 21.234)
```

#### Return a substring

You can get a substring from a character object using the `substr` function

```{r}
new_str <- 'How much wood would a woodchuck chuck?'
substr(new_str, start = 5, stop = 13)
```

#### Replacing characters in a string

You can replace characters in a string using the `sub` function

```{r}
sub('much', 'little', new_str)
# help(sub)
```

### Logical

A logical value is commonly referred to as a boolean value. It is the result of a comparison between two variables. It is represented as `TRUE` or `FALSE.`

```{r}
x <- 1
y <- 3
z <- x > y
z
```

\

```{r}
x <- TRUE
y <- FALSE
x & y
x & x
x | y

```

#### Logical Operators

The standard logical operations are:

-   `&` - And

-   `|` - Or

-   `!` - Not

```{r}
# See above
```

```{r}
# See above
```

```{r}
# See above
```

```{r}
# See above
```

### Complex

A complex object in R is created using the symbol `i` . The use of the i symbol is in reference to imaginary numbers.

```{r}
z <- 1 + 2i
z
```

```{r}
class(z)
```

## Comparison Operators

You can use the following operators to compare two variables and return a logical object:

-   `<` Less than

-   `>` Greater than

-   `<=` Less than or equal to

-   `>=` Greater than or equal to

-   `==` Is equal to

-   `!=` Is NOT equal to

```{r}
12 > 10
```

```{r}
12 < 10
```

```{r}
12 <= 10 

```

```{r}
12 >= 10
```

```{r}
12 == 10
```

```{r}
12 != 10
```

## Homework

1.  Create two strings with your first and last name, respectively. Then concatenate them, and print the new value as "My name is \<your newly created concatenated variable will go here\>"

    ```{r}
    first_name <- 'Scott'
    last_name <- 'McLaughlin'
    full_name <- paste(first_name,last_name)
    full_name
    ```

2.  What values can a BOOLEAN contain?

TRUE or FALSE\
